\documentclass[11pt]{article}
\usepackage{fullpage,amsmath,amssymb,amsfonts,url}

\title{
  CSCE 355 Programming Assignment
}
\author{Spring 2017}
\date{Due Mon, Apr 24, 2017 at 11:55 pm EST}


\begin{document}

\maketitle

For the programming portion of the course (10\% of your grade) you are
to write programs that perform two or more of a choice of six tasks
involving DFAs.  You may choose to implement two or more of the
following functions:
\begin{description}
  \item[Simulating a DFA\@.]
    After reading the description of a DFA $N$ from a text file, read
    any number of strings over $N$'s alphabet from another text file,
    indicating (to standard output) whether $N$ accepts or rejects
    each of the strings.
  \item[Minimizing a DFA\@.]
    Read the description of a DFA from a text file, and write (to
    standard output) the description of a minimal equivalent DFA\@.
  \item[Text search.]
    Read a string $w$ from a text file, and write to standard output
    the description of a DFA that accepts a string $x$ iff $w$ is a
    substring of $x$.
  \item[Closure properties: complement and intersection.]
    Read one or two descriptions of DFAs from text files (depending on
    the type of transformation asked for), and write to standard
    output a description of the transformed DFA---either the
    complementary DFA or the product construction.
  \item[Closure property: inverse homomorphic image.]
    Read from a text file the description of a DFA $A$ over some
    alphabet $\Gamma$ and (from another text file) the description of
    a homomorphism $h:\Sigma^*\rightarrow\Gamma^*$ (where $\Sigma$ is
    some alphabet), and write to standard output the description of a
    DFA recognizing $h^{-1}(L(A))$.
  \item[Determining properties of a DFA\@.]
    Read from a text file the description of a DFA $A$ and write to
    standard output (a) whether or not $L(A) = \emptyset$ and (b)
    whether or not $L(A)$ is finite.
\end{description}

\section{Details}

You are only required to implement your choice of \emph{two} of the
six alternatives above, but you may implement more than two if you
want a modest amount of extra credit (1\% of your final grade per
additional task correctly implemented).  As you decide which two,
notice that some combinations of tasks can share code.  For example,
all except the third require reading and internalizing a DFA
description; the second, third, fourth, and fifth require writing a
DFA description (see Section~\ref{sec:dfa-format}).

The grading of your work will be automated via scripts run on Linux
(These scripts are found from the project homepage:
\url{http://www.cse.sc.edu/~pade/csce355/}).  Therefore we are
requiring you to stick to a simple, uniform interface for your
program:  Your program must be able to be run via a simple
command-line invocation on a GNU/Linux box in the department (e.g.,
oak or ash), and all I/O will be ASCII text.  Input is read either
from text files (given as command line arguments) or standard input,
and output is written to standard output.  You may write your program
in any programming language you want, provided it is implemented on
the Linux machines in the CSCE lab.  We recommend Java or C or C++ or
Perl or Python or Ruby or ML or Haskell or \ldots.  To repeat: your
program (after compiling, if necessary) should be a stand-alone
executable that can be run directly from the Linux shell, requiring no
special user interface to execute (e.g., Eclipse).\footnote{It's OK if
  we need to invoke the JVM by preceding your main class name with
``java'' on the command line.}  More about this below.

This assignment is not meant to be overly taxing or time-consuming.
Most of the time spent will probably be to get the input read in
correctly.

\subsection{Tasks}

Here are more specific details for each of the six options above:

\subsubsection{Simulation}
First read the description of the DFA\@.  The description is in a file
named by the first command line argument (see
Section~\ref{sec:dfa-format} for information about the format of this
file).  Next read a series of zero or more strings from another text
file named by the second command line argument.  These are inputs to
the DFA\@.  Each string will take up an entire line of text, ending
with a newline character (which is not included in the string).  For
each string, write to standard output either ``accept'' or ``reject''
according to the behavior of the DFA on the string.  End each output
word with a newline.

\subsubsection{Minimization}
The input DFA is read from a text file given as the first command
line argument, and the output DFA is written to standard output.
You should use the table-of-distinguishable-states method
described in class to do the minimization.  You do not need to
output the distinguishability table, but you may optionally do so
to standard error if you want.  Both the input and output DFA
descriptions adhere to the format described below.  The ordering
of the states of the output DFA is not completely determined, so
correct answers may differ up to the ordering of states.  That's
OK\@; we'll take that into account using our own script that
determines whether two DFAs are the same up to reordering states.

\subsubsection{Text search}
This task is similar to Assignment 4, problem 1.  The name of the
file containing the text string is given as the sole command line
argument.  You should assume that the input string and the output
DFA are both over the alphabet consisting of the 26 lower case
letters \texttt{abcdefghijklmnopqrstuvwxyz} only.

\subsubsection{Bool Ops}
There will be either one or two command line arguments.  If one,
then it is the name of a file containing the description of a DFA
$A$, and you are to output the complementary DFA $\neg A$ (see
Definition~7.4 of the course notes).  If two arguments, then they
both are names of text files containing DFAs $A$ and $B$, and you
are to output the product construction $A\wedge B$ (see
Definition~7.4 of the course notes).  $A$ and $B$ will have the
same alphabet.  In each case, write to standard output.

\subsubsection{Inv. Hom. Image}
The first command line argument is the name of a text file containing
the description of the DFA, and the second argument gives a text file
containing the description of the homomorphism.  See
Section~\ref{sec:hom-format} for the description format for the
homomorphism.  Output is to standard output.

\subsubsection{Properties}
The sole command line argument gives the file describing the
DFA\@.  Output is to standard output and consists of just two
words separated by whitespace: the first is either
``\texttt{empty}'' or ``\texttt{nonempty}'', and the second is
either ``\texttt{finite}'' or ``\texttt{infinite}''.  Note that
the language of a DFA is nonempty iff some final state is
reachable from the start state, and the language is infinite iff
there is some cycle that is reachable from the start state and
from which a final state is reachable.

In all cases, you may assume that the files adhere to their respective
formats, i.e., you won't need to error-check an input file.  Note that
all input is from text files given as command line arguments; your
programs should not expect any input from standard input.  Your
programs' ``official'' output goes to standard output (not a text
file), but you may also print anything you want to standard error; the
grading program will ignore anything sent to standard
error.\footnote{We call the three I/O streams open by default on
GNU/Linux programs \emph{standard input} (buffered keyboard input by
default), \emph{standard output} (buffered screen output by default),
and \emph{standard error} (unbuffered output sent to the screen by
default, even if standard output is redirected by the system).  Some
programming environments may use different names for these streams,
e.g., C programs using stdio.h for high-level I/O call these stdin,
stdout, and stderr, respectively; C++ programs typically use cin,
cout, and cerr for the same purpose.}

Your code should be economically written, well-structured, and
well-commented, following the common stylistic guidelines of the
programming language you use.  The code should also be reasonably
efficient, but this is a secondary requirement.  If your code runs
correctly, we won't really look too closely at the source code.  If it
does not run correctly, however, the source code style is more
important.

\subsection{File Formats}

\subsubsection{DFA}\label{sec:dfa-format}

We will use a common ASCII text format for describing DFAs.  If you
read a DFA description as input, you can expect it to be in this
format, i.e., you don't need to error-check the description.  If you
write a DFA as output, you must use the same format.  (This allows the
flexibility of using the output of one program as the input to
another.)  The format is meant to be both simple to parse and easily
readable by humans.

Here is a sample description of a $5$-state DFA that accepts a binary
string iff the numerical value of the string differs by $1$ from a
multiple of $5$:
\begin{verbatim}
Number of states: 5
Accepting states: 1 4
Alphabet: 01
0 1
2 3
4 0
1 2
3 4
\end{verbatim}

Generally,
\begin{itemize}
  \item
    The first line starts with ``Number of states: '' followed by a
    single positive decimal integer giving the number of states of the
    DFA\@.  If the DFA has $n$ states, then we assume that the state
    set is $Q = \{0,1,2,\ldots, n-1\}$, with $0$ \emph{always} being
    the start state.
  \item
    The second line starts with ``Accepting states: '' followed by a
    list of nonnegative integers indicating the states that are
    accepting.  The numbers are separated by white space (strings of
    one or more space characters) and should appear in increasing
    order.
  \item
    The third line starts with ``Alphabet: '' followed by a single
    string of characters that takes up the rest of the line.  The
    characters in the string form the alphabet of the DFA\@.  They
    should come in order of increasing ASCII code, with no duplicates.
    We will restrict our alphabets to printable ASCII characters only,
    i.e., no control characters like carriage returns, line feeds,
    newlines, etc.  The space character is considered printable---and
    thus allowed---but the tab and DEL characters are not.  This means
    that the printable ASCII characters run from 32 through 126
    (decimal), inclusive.  The alphabet string starts with the first
    character after the space after the first colon, and runs through
    the end of the line (not including the final newline).
  \item
    The rest of the description consists of the transition table.  The
    rows of the table (each terminated with a newline character)
    correspond to the states of the DFA in numerical order.  Each row
    consists of a sequence of nonnegative integers separated by white
    space, one number for each alphabet symbol.  These numbers
    correspond to the destination states given the initial state and
    the alphabet symbol.  The order of entries within each row must
    correspond to the order of characters in the alphabet string given
    earlier.
\end{itemize}

Thus the sample DFA description above corresponds to the more
traditional tabular form
\begin{align*}
  \begin{array}{r||cc}
    & 0 & 1 \\
    \hline\hline
    \rightarrow 0 & 0 & 1 \\
    {*}1 & 2 & 3 \\
    2 & 4 & 0 \\
    3 & 1 & 2 \\
    {*}4 & 3 & 4
  \end{array}
\end{align*}

\subsubsection{Homomorphism}\label{sec:hom-format}

This only applies to the fifth task.  A homomorphism
$h:\Sigma^*\rightarrow\Gamma^*$ will be represented in a text file by
giving (a) the input alphabet $\Sigma$, (b) the output alphabet
$\Gamma$, and (c) the values $h(a)$ for all $a\in\Sigma$ in increasing
ASCII order.  For example, if $\Sigma = \{0,1,2\}$, \ $\Gamma =
\{a,b,c,d\}$, and $h(0) = abd$, $h(1) = \epsilon$, and $h(2) = acda$,
then the file will look like this:
\begin{verbatim}
Input alphabet: 012
Output alphabet: abcd
abd

acda
\end{verbatim}
(Notice the blank line denoting the empty string for $h(1)$.)  The
format of each of the two lines specifying the alphabets is similar to
the format of a DFA alphabet, above, with the same restrictions on the
alphabet.  Following the alphabet specifications is a series of
strings (one per line) giving the value of the homomorphism applied to
each of the input alphabet symbols, given in the same order as the
symbols in the input alphabet.

\subsubsection{Build/Run}\label{sec:build-format}

To score your files, you must tell the auto-grader how to build and
run your program for each task you have completed. Each build file
will be a standard ascii file, named \texttt{simulator.txt},
\texttt{minimizer.txt}, \texttt{searcher.txt}, \texttt{boolop.txt},
\texttt{invhom.txt}, or \texttt{properties.txt}, depending on the
corresponding task. The generic file format is:

\begin{verbatim}
Build:
  <list of commands to compile your program>
Run:
  <list of commands to run your program>
\end{verbatim}

\textbf{Do not forget to end your build file with an empty newline}

Suppose you implement a DFA simulator (first task, above) in Java, and
your main class is called \texttt{MySimulator}.  Then your submission
zip file should include a text file named \texttt{simulator.txt} whose
contents look like this:

\begin{verbatim}
# Lines like these are comments and will be ignored
Build:
  javac MySimulator.java
Run:
  java MySimulator
# Don't include command line arguments to the run command!
# The indenting is optional.
\end{verbatim}
For another example, suppose you implement a DFA minimizer in C as a
single compilation unit called \texttt{my\_minimizer.c}.  Then include
a file named \texttt{minimizer.txt} with something like the following
contents:

\begin{verbatim}
Build:
  gcc my_minimizer.c
  mv a.out my_minimizer
Run:
  ./my_minimizer
# Again, no command line arguments, please.  They will be supplied automatically.
\end{verbatim}

Note that you can have any number of build commands, and they will be
executed in order (in the directory containing your source files)
before the run command.  Always give the Build commands first before
the Run command.

Suppose instead that you have several compilation units for your
programs, including shared code, and a complicated build procedure,
but you have a single Makefile controlling it all, capable of
producing an executable called \texttt{my\_minimizer} and maybe other
executables as well.  Then the \texttt{minimizer.txt} file can just
look something like this:

\begin{verbatim}
Build:
  make my_minimizer
Run:
  ./my_minimizer
\end{verbatim}

and similarly for the other program(s) you implement.

As a final example, suppose you implement the inverse homomorphic
image closure property in Python, which is a scripting language that
can be run directly without a compilation step.  Then your
\texttt{invhom.txt} file might look like this:

\begin{verbatim}
Build:
Run:
  python my_inv_hom_maker.py
# You still need to say "Build:" even though there are no build commands.
\end{verbatim}

Finally, be sure your CSE Dropbox account exists and is accessible.
Do this early on to avoid last-minute glitches.

\subsection{Windows vs.\ Linux}

Windows-based text files end each line with the two-character sequence
\verb+\r\n+ (carriage return, newline), and GNU/linux/Mac OS X and
similar systems expect only an \verb+\n+ ending each line.  We
strongly recommend against doing your development on a Windows box,
but if you absolutely must, be aware that the files
\texttt{simulator.txt}, etc.\ may not work properly with the test
script if they are just copied over without newline conversion.  (This
is the cause of many mysterious failures when running the test
script.)  Our linux boxes support the \texttt{scc} command.  Running

\begin{verbatim}
    scc my_text_file.txt
\end{verbatim}

converts all \verb+\r\n+ sequences to \verb+\n+ in
\texttt{my\_text\_file.txt}.  (Note that this command alters the
contents of the file and does \emph{not} produce a backup copy.)


\section{Notes and hints}

One way to check if $L(A)=\emptyset$ for some DFA $A$ is to first
minimize it then check if the result is the $1$-state DFA with no
accepting states.  (More generally, an efficient way to check whether
two DFAs are equivalent is to minimize both of them, then check
whether the results are isomorphic (which is fairly quick, since there
can be at most one isomophism between any two sane DFAs).)  Minimizing
also helps for checking whether $L(A)$ is finite.  There can be at
most one dead state (from which no final state is reachable) in a
minimal DFA, so it it easy to isolate.  A minimal DFA recognizes an
infinite language if and only if it has a cycle not involving the dead
state.

Generally, you cannot assume any \textit{a priori} bounds on the
number of states of a DFA or the length of $h(a)$ where $h$ is a
homomorphism and $a$ is a symbol in the input alphabet of $h$.  In
other words, a DFA may less than $10$ states or more than $10,000$
states.  Similarly, the length of $h(a)$ may be $0$ or over $10,000$
So you will not be able to declare a buffer to hold $h(a)$ values
whose length is known at compile time.

\section{Testing and Grading}

As we mentioned, your project will be graded automatically.  We will
use the supplied testing program located on the course website and
test files in a test suite directory to test and grade your project.
\emph{All these files are available to you soon from the project
homepage,} so that you can see how your code will be tested and even
run the test program yourself to see in advance how well you do.  Just
to be perfectly clear: we will grade your project by running it
through an automated testing program.  We will not run your code
personally.  The comments produced by that program will determine your
grade.  This means that you will not get credit for attempting to do
something.  You will only get credit for what actually works, as
determined by the test's output.

Generally, your output does not have to match the solution output
exactly.  For example, if you minimize a DFA, your output (a minimal
DFA) only needs to match our solution up to relabeling (actually,
re-ordering) of states.  If two DFAs are the same up to
relabeling/re-ordering of states, we say that they are
\emph{isomorphic}.  The Myhill-Nerode theorem says that two minimal
DFAs are equivalent if and only if they are isomorphic; therefore, the
minimal DFA is unique \emph{up to isomorphism}.  We have a program
(that we also make available to you) that tests whether two DFAs are
isomorphic, and we will use this program to compare your output with
ours.  For another example, if you implement the fourth task, which
requires you to build a DFA for the intersection $L(A)\cap L(B)$ for
two given DFAs $A$ and $B$, we will only check that your output DFA is
\emph{equivalent} to the product construction, which suffices to
recognize $L(A)\cap L(B)$.  Since the product construction does this
already, the number of states of your output DFA should be \emph{no
more} than the product of the sizes of the state sets of $A$ and of
$B$.

\section{Submission}

Submission will be via CSE Departmental Dropbox (Moodle).  Upload a
single file, either a.zip file or a.tar.gz file, containing

\begin{enumerate}
  \item
    all your source code files, which should all be in the same
    directory, i.e., no subdirectories (and no automatically generated
    files, please),
  \item
    a README file indicating which two of the options above you are
    implementing for full credit, along with anything else you want to
    tell us (we will read this with our own eyes), and
  \item
    for each task you implement, a ``build/run'' text file giving
    commands to compile and/or run the program performing the task
    (see Section~\ref{sec:build-format} for more information).
\end{enumerate}


IMPORTANT NOTE:\@You \emph{must} use either the ZIP format (file
extension `.zip') or the GZIPPED TAR format (file extension `.tar.gz')
for your submission file.  Your file will be de-archived either with
\texttt{unzip} or with \texttt{gunzip; tar -xf}, depending on your
file name's extension.  Do not use any other archive format,
particularly the RAR format, which is proprietary to Windows (I
personally do not have Windows on any machine I use).  If you deviate
from the allowed formats, you risk getting zero credit for the entire
assignment.

\section{Do your own work}

The code you write and submit must be yours alone.  You may discuss
the homework with others at the conceptual level only, but you may not
copy code directly from any other source, even if you modify it
afterwards.  Likewise, you must take reasonable precautions not to let
your code be copied by anyone else.  Violating this policy constitutes
a violation of the Carolina Honor Code, and will have serious
consequences, including, but not limited to, failure of the course.

If you have any questions about what this policy means, please ask.

\end{document}
